\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[hidelinks]{hyperref}
\usepackage{longtable}
\usepackage{array}
\usepackage{enumitem}

\title{Build-a-Bot Stadium\\As-Built Summary}
\author{Project Snapshot}
\date{8 April 2026}

\begin{document}
\maketitle

\section{Purpose and Scope}
This document summarizes the current \emph{as-built} state of Build-a-Bot Stadium at a practical, implementation-aligned level. It is intended as a concise technical orientation for contributors and reviewers who need to understand what the platform currently does, how it is organized, and which constraints remain.

The summary focuses on implemented capabilities and operational characteristics rather than design aspirations.

\section{System Overview}
Build-a-Bot Stadium is currently implemented as a Go server runtime plus an in-repository Avalonia desktop client (Track A alpha).

Server runtime primary external surfaces:
\begin{itemize}[leftmargin=*]
  \item A TCP bot protocol endpoint (default \texttt{:8080})
  \item An HTTP dashboard and viewer endpoint (default \texttt{:3000})
\end{itemize}

Desktop client surface:
\begin{itemize}[leftmargin=*]
  \item Avalonia unified workspace shell in \texttt{bbs-client/} with left bot panel, center activity host, and right server panel
\end{itemize}

The platform follows a plugin-only game architecture:
\begin{itemize}[leftmargin=*]
  \item All games are process-based external plugins discovered at runtime
  \item Supports contract-compliant interaction patterns (multi-session, autonomous, or single-session)
  \item Plugins are optional; minimal deployment requires only the core server
\end{itemize}

As of v0.3.0, the runtime includes Stage 0 persistence and deployment-hardening features:
\begin{itemize}[leftmargin=*]
  \item SQLite-backed durable records for server identity, bot profiles, match history, and federation outbox
  \item Runtime path contract for templates, plugin directory, config path, and data path
  \item Linux FHS packaging profile with native Debian package support and systemd service integration
\end{itemize}

Primary references:
\begin{itemize}[leftmargin=*]
  \item \texttt{README.md}
  \item \texttt{docs/architecture/ARCHITECTURE.md}
\end{itemize}

\section{Implemented Feature Set}
\subsection{Arena Runtime and State Management}
The manager in \texttt{stadium/} remains the active in-process authority for sessions and arenas. Arena lifecycle, participant attachment, outcomes, and dashboard snapshot broadcasting are handled centrally.

Notable implemented behavior:
\begin{itemize}[leftmargin=*]
  \item Zero-player to multi-player arena activation support via game policy interfaces
  \item Move clock and handicap policy handling based on game capabilities
  \item Match archival with replay-oriented metadata
  \item Watchdog-based arena cleanup
\end{itemize}

Primary references:
\begin{itemize}[leftmargin=*]
  \item \texttt{docs/architecture/ARCHITECTURE.md}
  \item \texttt{stadium/manager.go} and related files in \texttt{stadium/}
  \item \texttt{stadium/manager\_persistence\_hooks.go}
  \item \texttt{stadium/persistence\_sqlite.go}
\end{itemize}

\subsection{Bot Command Protocol (TCP)}
A newline-delimited command protocol is exposed for bot clients. Core commands include registration, arena creation/joining, move submission, observation, and disconnect.

All command responses use a single JSON envelope shape (\texttt{status}, \texttt{type}, \texttt{payload}) over JSONL, including greeting and lifecycle responses.

Current protocol supports dynamic game arguments in \texttt{CREATE}, enabling consistent invocation across plugin games.

Primary references:
\begin{itemize}[leftmargin=*]
  \item \texttt{docs/reference/PROTOCOL.md}
  \item \texttt{cmd/bbs-server/main.go}
\end{itemize}

\subsection{Dashboard and Viewer Surfaces}
The dashboard provides operational visibility and control over runtime state, including owner-scoped and admin-scoped control paths.

Implemented dashboard capabilities include:
\begin{itemize}[leftmargin=*]
  \item Live state streaming via SSE and WebSocket
  \item Owner token flow (register bot, create/join/leave/eject)
  \item Admin control flow protected by \texttt{BBS\_DASHBOARD\_ADMIN\_KEY}
  \item Dynamic arena creation forms driven by live game catalog metadata
  \item Live and replay viewer pages with client-side rendering
  \item Plugin discovery diagnostics panel showing enabled state, scan directory, and per-manifest load/skip status
  \item Admin debug endpoints for server identity, outbox, and recent matches
\end{itemize}

\subsubsection{Viewer Architecture}
The viewer now uses a client-side rendering model where all game visualization logic runs in the browser:
\begin{itemize}[leftmargin=*]
  \item Server emits only raw game state frames (no structured viewer payloads)
  \item Each plugin declares optional \texttt{viewer\_client\_entry}: path to a JavaScript rendering bundle
  \item \texttt{/viewer/plugin-entry?game=<name>} serves the plugin's JS bundle to the browser
  \item Browser runtime loads the JS and calls the plugin's renderer with raw state
  \item Live and replay viewing use identical state streaming and rendering pipeline
\end{itemize}

This design ensures zero server-side render coupling, letting game developers own visualization entirely.

Primary references:
\begin{itemize}[leftmargin=*]
  \item \texttt{cmd/bbs-server/dashboard.go}
  \item \texttt{cmd/bbs-server/viewer.go}
  \item \texttt{cmd/bbs-server/templates/index.html}
  \item \texttt{cmd/bbs-server/templates/dashboard-state.html}
  \item \texttt{cmd/bbs-server/templates/viewer.html}
  \item \texttt{docs/architecture/ARCHITECTURE.md} (Viewer Architecture section)
  \item \texttt{docs/guides/DASHBOARD\_REMOTE\_QUICKSTART.md}
\end{itemize}

\subsection{Process-Based Game Plugin System}
A process plugin model is the only active game source:
\begin{itemize}[leftmargin=*]
  \item \texttt{BBS\_ENABLE\_GAME\_PLUGINS=true} activates plugin discovery
  \item \texttt{BBS\_GAME\_PLUGIN\_DIR=/path/to/plugins} (runtime-resolved defaults by mode)
\end{itemize}

Plugin manifests are discovered from \texttt{*.json} files in the configured directory. Plugin games communicate with the host over JSONL RPC via stdin/stdout. Manifests must declare \texttt{viewer\_client\_entry} to be valid and discoverable.

Implemented plugin author support:
\begin{itemize}[leftmargin=*]
  \item Shared protocol and helper runtime (\texttt{games/pluginapi})
  \item Reference plugin command (\texttt{cmd/bbs-game-counter-plugin})
  \item Dashboard-integrated arg schema rendering
  \item Discovery diagnostics in the dashboard
\end{itemize}

Primary references:
\begin{itemize}[leftmargin=*]
  \item \texttt{docs/architecture/ARCHITECTURE.md}
  \item \texttt{docs/guides/PLUGIN\_AUTHORING.md}
  \item \texttt{games/pluginapi/protocol.go}
  \item \texttt{games/pluginapi/server.go}
  \item \texttt{games/plugin\_host.go}
  \item \texttt{games/plugin\_process\_game.go}
\end{itemize}

\subsection{Local Bridge Agent for Bot Authors}
\texttt{bbs-agent} provides a local JSONL bridge abstraction for bot logic, separating local policy code from direct BBS TCP protocol handling.

Implemented bridge model:
\begin{itemize}[leftmargin=*]
  \item Local socket endpoint and message contract v0.2
  \item \texttt{hello} / \texttt{welcome} / \texttt{turn} / \texttt{action} message loop
  \item Support for domain-agnostic contract-compliant usage patterns
\end{itemize}

\subsection{Desktop Client Alpha (Track A)}
The repository now includes a functional desktop client alpha implementation in \texttt{bbs-client/}. Current implemented scope aligns with Track A slices A0-A3 baseline:
\begin{itemize}[leftmargin=*]
  \item Unified single-view workspace with collapsible left/right side panels and context-driven center host
  \item Durable local client state via SQLite (identity, bot profiles, known servers, plugin cache, runtime state contracts)
  \item First-launch client identity bootstrap with idempotent schema initialization/migration
  \item Bot registration/edit flow in center workspace with persistence and reload on restart
  \item Deploy-based runtime orchestration with persisted session state projection into bot card status text
  \item Server-context owner-token metadata resolved from selected known server profiles, with active-session arena dropdowns refreshed from the selected server's live arena list
  \item Iterated side-panel UX polish (card-centric selection/hover visuals, keyboard focus-visible cues, centered headers, icon collapse controls)
\end{itemize}

Current limitations relative to full Track A alpha intent:
\begin{itemize}[leftmargin=*]
  \item Bot glow-state color semantics (amber/green/red) remain pending as a dedicated follow-up slice
  \item Known-server persistence/probing, server detail workflows, and owner-token-gated actions remain planned work
\end{itemize}

Primary references:
\begin{itemize}[leftmargin=*]
  \item \texttt{bbs-client/README.md}
  \item \texttt{docs/releases/TRACK\_A\_TASK\_CHECKLIST.md}
  \item \texttt{docs/releases/ROADMAP.md}
\end{itemize}

Primary references:
\begin{itemize}[leftmargin=*]
  \item \texttt{docs/reference/BBS\_AGENT\_CONTRACT.md}
  \item \texttt{cmd/bbs-agent/README.md}
  \item \texttt{cmd/bbs-agent/main.go}
\end{itemize}

\subsection{Persistence and Federation Outbox (Stage 0)}
The runtime now includes a persistence boundary and SQLite implementation for durable operational state. The current Stage 0 persistence scope includes:
\begin{itemize}[leftmargin=*]
  \item Server identity and global registration support records
  \item Bot profiles and match history persistence
  \item Outbox records for asynchronous federation-style event publishing
  \item Receipt/dedupe support for mock federation ingestion flows
\end{itemize}

Outbox processing includes retry and backoff behavior through a dedicated worker loop and HTTP publisher abstraction.

Primary references:
\begin{itemize}[leftmargin=*]
  \item \texttt{stadium/persistence\_contracts.go}
  \item \texttt{stadium/persistence\_sqlite.go}
  \item \texttt{stadium/manager\_outbox\_worker.go}
  \item \texttt{stadium/outbox\_http\_publisher.go}
  \item \texttt{docs/architecture/ADR\_SQLITE\_STAGE0.md}
\end{itemize}

\subsection{Runtime Path Contract and Packaging Profile}
Runtime filesystem paths are now explicit and configurable, enabling predictable behavior across development and packaged deployments.

Implemented path/packaging behavior:
\begin{itemize}[leftmargin=*]
  \item Template path, plugin directory, config path, and data path resolved through runtime contract helpers
  \item Linux packaging profile for FHS-compatible deployment layout
  \item Debian package scaffolding with install scripts and systemd unit
  \item Build automation for \texttt{.deb} artifacts through repository scripts and Makefile target
\end{itemize}

Primary references:
\begin{itemize}[leftmargin=*]
  \item \texttt{cmd/bbs-server/runtime\_paths.go}
  \item \texttt{docs/architecture/ADR\_RUNTIME\_PATHS\_STAGE0.md}
  \item \texttt{docs/deployment/LINUX\_PACKAGING\_PROFILE.md}
  \item \texttt{debian/}
  \item \texttt{scripts/build-deb.sh}
  \item \texttt{Makefile}
\end{itemize}

\subsection{Packaging, Deployment, and Release Automation}
Deployment guidance exists for both Docker and non-Docker TrueNAS workflows, and release automation publishes Linux binaries for amd64 and arm64.

Primary references:
\begin{itemize}[leftmargin=*]
  \item \texttt{deploy/truenas/README.md}
  \item \texttt{deploy/truenas-no-docker/README.md}
  \item \texttt{.github/workflows/release-bbs-server.yml}
\end{itemize}

\subsection{Validation and CI Status}
A dedicated manifest-validation command and workflow are present:
\begin{itemize}[leftmargin=*]
  \item \texttt{cmd/bbs-plugin-manifest-lint/main.go}
  \item \texttt{.github/workflows/validate-plugin-manifests.yml}
\end{itemize}

This improves plugin packaging hygiene by validating schema shape, field constraints, duplicates, and protocol compatibility.

The repository also contains substantial Go test coverage across server, stadium manager, plugin host, and runtime-path/persistence components.

\section{Current Gaps and Shortcomings}
The following are the most immediate gaps visible in the current implementation snapshot.

\subsection{Durability and Data Lifecycle}
\begin{itemize}[leftmargin=*]
  \item Persistence is currently Stage 0 and intentionally scoped; not all live runtime state is durably modeled.
  \item Active arena/session continuity across restart is still limited and should be treated as best-effort rather than full crash-recovery.
\end{itemize}

Implication: durability has materially improved, but full operational resilience still requires further lifecycle and recovery work.

\subsection{Security and Trust Model}
\begin{itemize}[leftmargin=*]
  \item Bot channel remains plain TCP without built-in transport security.
  \item Identity and owner-token controls are bearer-style.
  \item Plugin trust assumes local trusted filesystem and trusted binaries.
\end{itemize}

Implication: deployment in untrusted networks requires external controls (for example TLS termination, network segmentation, host hardening).

\subsection{Scalability and Concurrency Model}
\begin{itemize}[leftmargin=*]
  \item Concurrency is now split across two levels: a manager-level mutex for global registries/counters/subscribers and a per-arena mutex for arena-local mutable state (players, observers, status, clocks, move history).
  \item Many hot-path arena operations lock manager state briefly to resolve arena/session references, then transition to arena-local locking for state mutation.
  \item Some global workflows (for example, snapshot/list assembly and watchdog scans) still hold manager-level coordination while iterating arenas, which can introduce contention under higher load.
\end{itemize}

Implication: the model is no longer purely coarse-grained and already localizes part of contention to individual arenas, but manager-level coordination remains a scaling pressure point for large concurrent populations.

\subsection{Plugin Governance and Isolation}
\begin{itemize}[leftmargin=*]
  \item Plugin discovery and shape validation exist, but stronger controls are not yet present (for example plugin signing, attestation, quotas, sandbox policies).
  \item Health enforcement and richer lifecycle supervision are limited.
\end{itemize}

Implication: plugin extensibility is practical now, but governance and safety controls are early-stage.

\subsection{Automated Test Coverage}
Automated test coverage is now present across key modules, but depth and scenario realism can still be expanded (especially recovery/failure-path and higher-load behaviors).

Implication: baseline regression guardrails are in place; next gains come from broader integration and fault-injection style coverage.

\section{Suggested Near-Term Focus}
A pragmatic next phase that aligns with the current architecture would be:
\begin{enumerate}[leftmargin=*]
  \item Expand Stage 0 persistence into stronger restart/recovery semantics for active arenas/sessions.
  \item Increase security hardening for production-like deployments (transport protection, token handling improvements, plugin controls).
  \item Expand plugin operational controls (health checks, richer diagnostics, controlled resource limits, and isolation posture).
  \item Add deeper integration and failure-path testing around persistence, outbox delivery, and packaging/runtime mode transitions.
\end{enumerate}

\section{Document Map (Primary Reading Order)}
\begin{longtable}{|p{0.34\textwidth}|p{0.60\textwidth}|}
\hline
\textbf{Document} & \textbf{Primary Use} \\
\hline
\path{README.md} & High-level platform orientation, run instructions, plugin quickstart, and usage examples. \\
\hline
\path{docs/architecture/ARCHITECTURE.md} & Runtime decomposition, component responsibilities, plugin model, and known boundaries. \\
\hline
\path{docs/reference/PROTOCOL.md} & TCP command protocol and response conventions for bot clients. \\
\hline
\path{docs/reference/BBS_AGENT_CONTRACT.md} & Local bridge JSONL contract for \texttt{bbs-agent} integrations. \\
\hline
\path{docs/deployment/LINUX_PACKAGING_PROFILE.md} & Linux FHS packaging contract, \texttt{.deb} build path, and packaged runtime layout. \\
\hline
\path{docs/architecture/ADR_RUNTIME_PATHS_STAGE0.md} & Runtime path contract decisions for templates/plugins/config/data resolution. \\
\hline
\path{docs/architecture/ADR_SQLITE_STAGE0.md} & Stage 0 SQLite persistence architecture and tradeoffs. \\
\hline
\path{docs/guides/PLUGIN_AUTHORING.md} & Plugin manifest schema, local workflow, CI checks, and release checklist. \\
\hline
\path{docs/guides/DASHBOARD_REMOTE_QUICKSTART.md} & Remote operator workflow for connecting bots through dashboard-issued owner tokens. \\
\hline
\path{bbs-client/README.md} & Desktop client build/run guide and current Track A alpha surface description. \\
\hline
\path{docs/releases/TRACK_A_TASK_CHECKLIST.md} & Track A implementation checklist with per-slice status and completion criteria. \\
\hline
\path{deploy/truenas/README.md} & Docker-based TrueNAS deployment and update operations. \\
\hline
\path{deploy/truenas-no-docker/README.md} & Non-Docker TrueNAS deployment and update operations. \\
\hline
\path{.github/workflows/release-bbs-server.yml} & Binary release automation (linux amd64 and arm64 artifacts). \\
\hline
\path{.github/workflows/validate-plugin-manifests.yml} & CI manifest validation workflow for plugin packaging quality. \\
\hline
\end{longtable}

\section{As-Built Conclusion}
Build-a-Bot Stadium has progressed from a narrow game host into a fully plugin-driven, extensible bot-arena platform with:
\begin{itemize}[leftmargin=*]
  \item plugin-only game discovery and runtime,
  \item client-side rendering model for game visualization,
  \item multi-mode arena semantics supporting domain-agnostic interaction patterns,
  \item functional process-plugin architecture with manifest validation,
  \item Stage 0 durable persistence and outbox infrastructure,
  \item runtime path contracts supporting both dev and packaged deployments,
  \item integrated dashboard-driven operations with owner and admin workflows, and
  \item practical deployment and release pathways.
\end{itemize}

The core extension and operator workflows are now demonstrably in place, with initial durability and packaging foundations implemented. The most significant remaining work is concentrated in stronger recovery semantics, security hardening, and broader operational test depth.

\end{document}
